Racial and Ethnic Differences in ETI

### 2023-08-04 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti.qmd
tidyverse formatting

```{r data cleaning}
#| include: false
source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

@@

```{r table 1}
#| label: tbl-1
#| tbl-cap: Race by year
annual %>%
  select(ReviewYear, race, Hispanicrace) %>%
  tbl_summary(by = ReviewYear, missing_text = "Missing")
```

```{r table 2}
#| label: tbl-2
#| tbl-cap: Race and ethnicity
demo %>%
  select(race, Hispanicrace) %>%
  tbl_summary(by = Hispanicrace) %>%
  modify_header(label = paste0("**", label(annual$Hispanicrace), "**")) %>%
  add_overall()
```

@@

@@

```{r}
#| label: tbl-3
#| tbl-cap: Characteristics by race
demo %>%
  select(race) %>%
  tbl_summary()
```

# Figure

>>>>

```{r data cleaning}
#| include: false
# source("/Users/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_race_eti_analysis_dataset.R")
load("/Volumes/Work/Vigers/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/analysis_dataset.RData")
```

@@

## Race by year

```{r t1}
annual %>%
  select(ReviewYear, race, Hispanicrace) %>%
  tbl_summary(
    by = ReviewYear, missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  modify_header(label = paste0("**Year**")) %>%
  bold_labels()
```

## Race and ethnicity overall

```{r t2}
demo %>%
  select(race, Hispanicrace) %>%
  tbl_summary(
    by = Hispanicrace,
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  modify_header(label = paste0("**", label(annual$Hispanicrace), "**")) %>%
  add_overall() %>%
  bold_labels()
```

@@

## Race percentages overall

```{r t3}
demo %>%
  select(race) %>%
  tbl_summary(statistic = list(all_continuous() ~ "{mean} ({sd})")) %>%
  modify_header(label = paste0("**Race**")) %>%
  bold_labels()
```

@@

## Characteristics by race

```{r t4}
demo %>%
  select(race, Gender, WasGenotyped, F508) %>%
  tbl_summary(
    by = race, missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  modify_header(label = paste0("**Race**")) %>%
  add_overall() %>%
  bold_labels()
```

```{r summarise by year}
#| cache: true
t5 <- encounter %>%
  group_by(eDWID, reviewyear) %>%
  summarise(across(c(race, Gender, WasGenotyped, F508, contains("insurance")), first),
    `Age at Last Visit` = last(na.omit(encounterage)),
    `Mean ppFEV` = mean(GLI_FEV1_pct_predicted, na.rm = T),
    `Weight at Last Visit` = last(na.omit(weight)),
    `Weight %ile at Last Visit` = last(na.omit(weightpercentile)),
    `BMI at Last Visit` = last(na.omit(bmivalue)),
    `BMI %ile at Last Visit` = last(na.omit(bmipercentile)),
    .groups = "drop"
  )
label(t5)[colnames(t5) %in% names(labels)] <- labels[match(colnames(t5), names(labels), nomatch = 0)]
```

## Characteristics by year

```{r t5}
#| message: false
t5 %>%
  select(
    -eDWID, reviewyear, race, Gender, WasGenotyped, F508, contains("insurance"),
    `Age at Last Visit`, `Mean ppFEV`, `Weight at Last Visit`,
    `Weight %ile at Last Visit`, `BMI at Last Visit`,
    `BMI %ile at Last Visit`
  ) %>%
  tbl_summary(
    by = reviewyear, missing_text = "Missing",
    statistic = list(all_continuous() ~ "{mean} ({sd})")
  ) %>%
  modify_header(label = paste0("**Year**")) %>%
  bold_labels()
```
